---
title: "p8105_hw3_cac2225"
author: "Courtney Chan"
date: "October 8, 2018"
output: github_document
---

##change the names, labels for each plot, rename for variables?
## interpret for each plot
## make sure they knit

```{r global settings for document}

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```

#Homework 3

##Problem 1

The dataset and library packages are first loaded.

```{r loading BRSS data}

library(p8105.datasets)
brfss_smart2010 = brfss_smart2010

```

```{r upload necessary packages}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(forcats)
library(ggridges)

```

To clean the dataset, the variable names will be cleaned using the janitor package, function clean_names.
The data is filtered based on topic, Overall Health, and responses Excellent to Poor. The response variable will be turned into a factor variable and its responses ordered.

```{r data cleaning step}
brfss_smart2010 = janitor::clean_names(brfss_smart2010)

filtered_brfss_smart2010 = filter(brfss_smart2010, topic == "Overall Health") %>% 
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>% 
  mutate(response = factor(response, levels = str_c(c("Excellent", "Very good", "Good", "Fair", "Poor"))))

```


```{r states observed at 7 locations in 2002}
filtered_brfss_smart2010 %>% 
  filter(year == "2002") %>% 
  distinct(locationdesc) %>% 
  separate(locationdesc, into = c("state", "county"), sep = "-") %>%   
  count(state) %>% 
  filter(n == 7)
```

To determine how many states were observed at 7 county locations, the number of distinct state-county values were determined, these values were seperated, and the number of times each state occurred was counted. The states that were counted as having appeared 7 times were filtered out.
The states are as follows: CT, FL and NC.

```{r creating spaghetti plot number locations in each state, 2002 to 2010}
filtered_brfss_smart2010 %>% 
  summarize(max_year = max(year), min_year = (min(year)))

plot_spaghetti = filtered_brfss_smart2010 %>% 
  separate(locationdesc, into = c("state", "county"), sep = "-") %>% 
  group_by(year, state) %>% 
  summarize(n_counties = n())

ggplot(plot_spaghetti, aes(x = year, y = n_counties, color = state)) + 
  geom_line() + 
  labs(
    title = "Spaghetti Plot Number of Locations in Each State, from 2002 to 2010",
    y = "Number of Locations"
  )
```

First I determined what the minimum and maximum year values are, which are 2002 and 2010, thus there was no need to filter by year any further. I created a new set of data, named plot_spaghetti, seperating the variable locationdesc to ultimately count the number of times a state is recorded within a certain year. I plotted this setting year on the x axis, number of times a state has appeared which equals number of locations per state, and a different color for each state.

```{r table of 2002, 2006 and 2010 proportion of Excellent responses in NY}
table_filt_brss_smart2010 = filtered_brfss_smart2010 %>% 
  filter(year %in% c("2002", "2006", "2010")) %>% 
  filter(response == "Excellent") %>% 
  filter(locationabbr == "NY") %>% 
  group_by(year, response, locationabbr) %>% 
  summarize(mean_Excellent = mean(data_value),
            sd_Excellent = sd(data_value))
```  

Dataset was filtered by year (2002, 2006 and 2010), response (Excellent) and locationabbr (NY). The mean and standard deviation of the proportion of Excellent responses across NY state were calculated using the group by and summarize functions.

```{r avg proportion of each response per state}

plot_avg_response = filtered_brfss_smart2010 %>% 
  group_by(year, locationabbr, response) %>% 
  summarize(proportion_response = mean(data_value)) %>% 
  ggplot(aes(x = year, y = proportion_response)) + 
  facet_grid(. ~ response) + 
  geom_violin() + 
  labs(
    title = "Five Panel Plot per Response type - distribution of State-level averages over time",
    y = "Proportion of each Response"
  ) + 
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 1), text = element_text(size = 7))

```

The mean values of the proportion of each response was generated using group by and summarize functions, grouped by year and response type. Using facet, a five panel plot per response type was created, each panel contained a violin plot of proportion of responses over the years. Thus over the years, the majority of the proportion of excellent responses were mainly around 20. The max proportion was about 30 and min proportion was about 12 in 2006. Over the years, the majority of the proportion of very good responses were around 35, with max proportion at around 43 and min proportion at around 26 in 2006. The majority of the proportion of good responses were around 30, with max proportion at 36 and min at 28 in 2006. The majority of the proportion of fair responses were around 10, with max at 27, min at 7 in 2006. The majority of the proportion of poor responses were around 4, with min at 3, max at 9 in 2006.

##Problem 2

First the data Instacart is loaded and the package forcats is loaded to be used for later on in the problem.

```{rloading Instacart data}

instacart = instacart

```

## Preliminary description of dataset

```{r Preliminary review of the data}

names(instacart)

```

To perform a preliminary review of the dataset, the names of the variables are first listed, by calling the names.

It seems that there are orders that are made with descriptions of the orders using variables such as giving each order an ID number (order_id), listing the day (order_dow) and hour (order_hour_of_day) the order was made, days since the prior order was made (days_since_prior_order) etc. Order_dow has values from 0 to 6, each number corresponding to a day of the week. The orders consist of products which are assigned product ids (product_id), described by what ailse and department they came from (aisle, aisle_id, department, department_id), ordered by which user (user_id), and the product's name (product_name). Key variables would include order_id, user_id, product_id, department_id and aisle_id, since by having these variables, one can determine important information, such as what product was ordered by whom and where that product came from. The data structure appears tidy since the variables are listed as column names and corresponding values are listed under each variable.

To describe the size of the dataset, the number of observations and number of variables are `r dim(instacart)`.

##Answering questions about the dataset instacart

```{r # of aisles and aisles with most orders}

instacart %>% 
  count(n_distinct(aisle_id))

instacart %>% 
    group_by(aisle, order_id) %>%  
    summarise(order_count = n()) %>% 
    group_by(aisle) %>% 
    summarise(total_order_count =
                sum(order_count)) %>%
  arrange(desc(total_order_count))

```

Using the n_distinct and count functions, there are 134 distinct aisle_ids, thus there are 134 aisles. Using the group_by and summarize functions, the number of orders made per aisle is counted and designated as order_count. Then for each aisle, the number of orders made is summed and arranged in descending order. The top 3 aisles that received the most orders are "fresh vegetables" at 150609 orders, "fresh fruits" with 150473 orders and "packaged vegetable fruits" with 78493 orders.

```{r number ordered items per aisle}

orders_count = instacart %>% 
    group_by(aisle, aisle_id, department) %>%
    summarise(total_order_count = n())

ggplot(orders_count, aes(fct_reorder(aisle, desc(total_order_count)), total_order_count, fill = department)) + 
  geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), text = element_text(size = 7))

```

A subset of data is generated, counting the total number of items ordered from each aisle, using the group_by and summarise functions. Using ggplot, the subset of data is plotted into a bar chart, with the aisles on the x axis and the total orders count on the y axis. The aisles are ordered and grouped by department id, with different colors representing different departments.

```{r most popular item in certain aisles}

most_popular_products = instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) %>% 
    group_by(aisle, product_name) %>%
    summarise(product_count = n()) %>% 
  filter(product_count == max(product_count))

```

To determine the most popular products in the three specific aisles, the dataset was first filterd by the three specific aisles. Then using the group_by and summarise functions, the number of products bought from each aisle were counted up and the product with the largest count were extracted. The most popular item from the baking ingredients aisle is Light Brown Sugar with 499 units ordered. The most popular item from the dog food care aisle is snack sticks chicken and rice dog treats with 30 units ordered. The most popular item from the packaged vegetables and fruits aisle is organic baby spinach with 9784 units ordered.

```{r Pink lady apples and coffee ice cream}

instacart %>% 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>% 
  group_by(order_dow, product_name) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hour) %>% 
  knitr::kable(digits = 1)

```

A 2x7 table is formed from the information filtered and summarized from the original dataset. Filter was used to extract information on the products pink lady apples and coffee ice cream. Group_by and summarise were used to calculate the mean hour of the day that which these products are order, by day of the week. Spread and knit were used to form the final 2x7 table. Looking at the table, for Coffee Ice Cream, the earliest mean time of the day the product is ordered is on Fridays (day 5) at 12.3 The latest time it's ordered is on Tuesday (day 2) at 15.4. For Pink Lady Apples, the earliest this product is ordered is on average Mondays (day1) at 11.4 and latest on average on Sundays at 13.4.

#Problem 3

The dataset is first loaded.

```{r load NY NOAA data}

ny_noaa = ny_noaa

```

The ny_noaa dataset consists of seven variables that describe weather station (id), amount of precipitation (prcp), amount of snowfall and depth (snow and snwd), temperature (tmax and tmin) and date (date). The total number of observation rows and number of columns are `r dim(ny_noaa)`. Key variables would be id and date, since these provide specific identification for each observation. The dataset is tidy since each variable is its own seperate column. The date column could be further seperated into month, year and day.

Each weather station might only collect a certain subset of variables, thus there is missing data. Missing data could cause an issue when we looking for trends, skewing or hiding any potential relationships that should normally appear in the data/plot visualizations.

```{r variables for y,m,d, units}

ny_noaa = ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-" ) %>% 
  mutate(prcp = as.numeric(prcp)/10,
         tmin = as.numeric(tmin)/10,
         tmax = as.numeric(tmax)/10)

```

Since the values of prcp, tmin and tmax are originally in tenths of a millimeter, they should first be converted into numeric variables, then divided by 10 to convert into millimeters. Additionally, to seperate the date variable, the separate function is used and the values are split via the symbol "-".

```{r most commonly observed values}

ny_noaa %>% 
  group_by(snow) %>% 
  summarise(common_snowfall = n()) %>% 
  arrange(desc(common_snowfall))

```

Using the group by and summarise functions, the most commonly observed values for snowfall are 0, NA and 25 mm. 0 may be the most common since it only really snows in the winter months in New York, which is three to four months out of the year. NA is very common since not all of the stations collect values on snowfall.

```{r plot mean max temps January and July}

plot_mean_max_temp = ny_noaa %>% 
  filter(month == "01" | month == "07") %>%
  group_by(id, month, year) %>% 
  summarise(mean_max_temp = mean(tmax)) %>% 
  ggplot(aes(x = id, y = mean_max_temp)) + 
  geom_point() + 
  facet_grid(. ~ month)

```

Looking at the scatterplot produced, the cluster of points only occurs amongst a certain subset of stations. Thus confirming that only certain stations collect certain datapoints/information - in this case only certain stations collected termperature while the rest had missing data. In January, the mean max temperatures are all in the lower range, varying from -10 to 10 degrees Celsius. There are a few outliers with temperatures lower than -10 degrees. In July, the mean max temperatures are all in the higher range, varying from 20 to 30 degrees, with a few outliers below 20 degrees.

```{r plot tmax vs t min}

plot_temp = ny_noaa %>% ggplot(aes(x = tmin, y = tmax)) + geom_smooth(se = FALSE)

plot_snow = ny_noaa %>% 
  filter(snow > 0, snow < 100) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_violin()

(plot_temp)/(plot_snow)

```

I created a two panel plot with a linear graph of tmin versus tmax on top and a violin plot of snowfall distribution greater than 0 mm and less than 100 mm for each year on the bottom panel. The linear plot shows that as t min increases, so does t max, as expected, since in the warmer months, the minimum and maximum temperatures are expected to increase. While in the colder months, both tmax and tmin would decrease. For each year, the distribution of snowfall seems similar. Most of the snowfall received occurs between 0 to 25 mm. There are measurements of snowfall clustering around 50 and 75 as well, while the larger amounts of snowfall taper off at 100 mm.